Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-07 thesis/paper.tex
review threats

%Description of inspection
From the APIs we collected method names, parameters and return types. (!!!Add more here. Maybe put it in a table?) The data were collected from both versions of the APIs, with the changes between the two versions being of particular interest. <-- do we still want to mention exactly what kinds of data we looked for? i.e. return types, enums, typedefs etc etc.

The APIs were inspected to determine where the changes had occurred. We used the Git Diff tool to find differences between two versions of files that existed in both versions of the API. Once we had identified the changes, we inspected the documentation and the code itself to determine if there was a reason for the change, for example, if a given function had been labelled deprecated between versions. Similar changes will be grouped together based on commonalities between them. (!!!reword nicely)

@@

%Interviews !!!improve later
After the initial source code inspection had been performed, interviews were conducted. The main purpose of the interviews was to validate our findings during the code inspection with the designers of the APIs. Furthermore, the interviews were conducted to provide additional insight into decisions and motives behind changes that might have been overlooked during the code inspection. The final reason for interviewing the designers of the APIs was to provide further insight of the users of the API and how their decisions, needs and requirements have affected the development of the API.

\subsection{Data Analysis} \label{data_analysis}
%Intro and summary
The data analysis was conducted in two major phases \cite{andersson2007spiral}. During the first phase, a hypothesis generating approach \cite{seaman1999qualitative} was used to be able to fulfil the exploratory goal of the study. It was during this phase that the analysis of the API code and documentation was carried out. The second phase included analysis of the interviews, the goal of which was to confirm the hypotheses \cite{seaman1999qualitative} that were formulated in the first phase. Both phases were performed iteratively, together with the data collection. This created the opportunity to improve the process of analysis as the study progressed, as well as allowing the study to adapt to possible changes in direction, due to the hypothesis generating approach of the first phase \cite{andersson2007spiral}. The analysis of the code and documentation was performed using grounded theory analysis \cite{seaman1999qualitative}, as it is recommended for hypothesis generating studies \cite{runeson2009guidelines} \cite{seaman1999qualitative}.

%Description of code analysis
More specifically, the approach undertaken was divided into three steps where the changes were first categorised based by following the decision tree for types \cite{chapin2001types}. The changes were then mapped according to the laws presented by \cite{lehman1980programs}. The changes were then analysed from a high-level point of view, the main goal of which was to identify trends in the changes. The final step involved analysis of both the trends and the changes. During this stage the hypotheses were generated and formulated. The same approach was taken to reduce the risks of research bias and human error in the data collection was also performed during results analysis.

%Description of interview analysis
Following the analysis of the archival data, the results of the interviews were analysed. The interviews were analysed using the same approach taken for the archival data. The results of this analysis was then compared and correlated with the results of the code analysis, with the goal of confirming the hypotheses that had been formulated. This support also supported triangulation of our results.

@@

Each change between versions has been identified and grouped into four particular categories reflecting what kind of change they were; New, Modified, Removed and General Trends.

Each section introduces the change and then discusses what the changes entailed, how we perceived them in relation to the twelve maintenance types, what our interviewee said about the changes and finally how the interviewees views of the changes relate to the maintenance categories. Each change was given a main category of maintenance and a subcategory if appropriate. This information will be used to answer \textbf{RQ1} and \textbf{RQ1.1}.
%This paragraph should be moved to Methodology

@@

\noindent
\textbf{Construct validity  } Threats to construct validity have largely been mitigated by structuring the results the study around already established and accepted theories \cite{chapin2001types} \cite{lehman1980programs}. Due to restricted access to the company's source code, we were not able to include versions previous to version 1.4 in our analysis. This means that the evolution of the API was studied during a rather short part of its lifetime ( "rephrase"). It was also not possible to analyse the implementation of the API, something which might have contributed to more accurate predictions of the motives behind the changes. We consider this to be remaining threats to construct validity.

\smallskip \noindent
\textbf{Internal validity  } Internal validity has been acknowledged by analysing the the code and documentation jointly. This has contributed to triangulating the results and discovering possible inconsistencies. Further triangulation of the results was performed by interviewing the API architect. Business-related factors might have had an impact on the evolution of the API, and these were not closely investigated. This was not in the scope of the study and would best be investigated in a study complementary to ours. The study revealed that the API users were significant to the evolution of the API. Since API users were not interviewed, it might be a affecting factor which was left unexplored.

\smallskip \noindent
\textbf{External validity  } The results of our study should be highly generalisable similar studies conducted on any type of APIs, since the evolution of the API has largely been driven by software-related factors. Our study should be of special interest to researchers studying platform or embedded APIs. The fact that the results of the study are presented in a structured way, by relating them to existing software evolution theories, should further increase the generalisability of our study.

>>>>

%Diffs / Excel documents - This was data collection

%Description of inspection
%Description of code analysis - Needs more emphasis on the qualitative grouping

%Identify the changes
The analysis of the code and documentation was performed using grounded theory analysis \cite{seaman1999qualitative}, as it is recommended for hypothesis generating studies \cite{runeson2009guidelines} \cite{seaman1999qualitative}.

The API code was inspected sequentially, starting with version 1.4. Both versions of the API underwent the same inspection. This involved extracting methods names, enums, data structures which were then subsequently stored in excel spreadsheets. Crosscase analysis was then performed on both sets of data retrieved from different versions of the API \cite{seaman1999qualitative}. Changes were grouped into categories that best represented what the change achieved and were added to the data tables (``Use excel? Is it ok?''). For example, additions, modifications and removals. After a change was found in the code, we found related documentation changes, or explanations that were relevant to that change and added them along side their corresponding changes in the excel sheets.

@@

%Interviews !!!improve later
After the initial source code inspection and preliminary analysis had been performed, interviews were conducted. The main purpose of the interviews was to validate our findings during the code inspection with the designers of the APIs. Ideally, the interviews would provide additional insight into decisions and motives behind changes that were not evident from the code inspection. The final reason for interviewing the designers of the APIs was to provide further insight of the users of the API and how their decisions, needs and requirements have affected the development of the API.

\subsection{Data Analysis} \label{data_analysis}
%Intro and summary
The data analysis was conducted in two major phases \cite{andersson2007spiral}. During the first phase, a hypothesis generating approach \cite{seaman1999qualitative} was used to be able to fulfil the exploratory goal of the study. It was during this phase that the analysis of the API code and documentation was carried out. The second phase included analysis of the interviews, the goal of which was to confirm the hypotheses \cite{seaman1999qualitative} that were formulated in the first phase. Both phases were performed iteratively, together with the data collection. This created the opportunity to improve the process of analysis as the study progressed, as well as allowing the study to adapt to possible changes in direction, due to the hypothesis generating approach of the first phase \cite{andersson2007spiral}.

%Initial state with the data - It's in tables as raw changes here
Once the preliminary groups of changes were identified, we had to identify what was different between the two versions.

%Grouped into the broader changes by now, new, added, modified and removed.
Each change between versions has been identified and grouped into four particular categories reflecting what kind of change they were; New, Modified, Removed and General Trends.

%Determining what kind of change it was
To determine if a change was new, we checked if it had existed in version 1.4 and then existed in version 2.0. For those files that did exist, we inspected using the Git Diff tool. This allowed us to find differences between files to determine if anything was changed. If we found a difference between two files where a function was removed, then it would be categorised as accordingly. Finally, for finding removals, we check if any files existed in version 1.4 and were no longer present in version 2.0.

%Documentation's role & trends
During the data collection stage, appropriate documentation was collected along side each change and was used to also determine what kind of change had occurred. Specifically, the documentation included a list of deprecated functions and files which allowed us to identify many changes easily. A high level analysis of the changes allowed us to identify general trends of the evolution.

%We have all the changes and our suggested theories. Categorise them now
Changes identified at this point were further categorised by following the decision tree for types \cite{chapin2001types}. Each change was given a main category of maintenance and a subcategory if appropriate. This information will be used to answer \textbf{RQ1} and \textbf{RQ1.1}.

%Description of interview analysis
The results of the interviews were analysed and categorised in the same way that our identified changes had been. The actual categorisations of the were tabulated and compared with our initial findings.

%Lehman's
The final step involved analysis of both the trends and the changes, combined with information gathered during the interview. These trends, combined with the information gathered in the interviews, allowed us to determine if Lehman's laws are applicable to APIs and therefore the answer to \textbf{RQ2} \cite{lehman1980programs}.

@@

%Where does this go?
Each section introduces the change and then discusses what the changes entailed, how we perceived them in relation to the twelve maintenance types, what our interviewee said about the changes and finally how the interviewees views of the changes relate to the maintenance categories. 
%This paragraph should be moved to Methodology

@@

\smallskip \noindent
\textbf{Construct validity  } Threats to construct validity have largely been mitigated by structuring the results of the study around already established and accepted software evolution theories \cite{chapin2001types} \cite{lehman1980programs}. Due to restricted access to the company's source code, we were not able to include versions previous to version 1.4 in our analysis. This means that the evolution of the API was studied during a rather short part of its lifetime ( "rephrase"). It was also not possible to analyse the implementation of the API, something which might have contributed to more accurate predictions of the motives behind the changes. We consider this to be remaining threats to construct validity.

\smallskip \noindent
\textbf{Internal validity  } Internal validity has been acknowledged by analysing the code and documentation jointly. This has contributed to triangulating the results and discovering possible inconsistencies. Further triangulation of the results was performed by interviewing the API architect. Business-related factors might have had an impact on the evolution of the API, and these were not closely investigated. This was not in the scope of the study and would best be investigated in a study complementary to ours. The study revealed that the API users were significant to the evolution of the API. Since API users were not interviewed, it might be an affecting factor which was left unexplored.

\smallskip \noindent
\textbf{External validity  } The results of our study should be highly generalisable to similar studies conducted on any type of APIs, since the evolution of the API has largely been driven by software-related factors. Our study should be of special interest to researchers studying platform or embedded APIs. The fact that the results of the study are presented in a structured way, by relating them to existing software evolution theories, should further increase the generalisability of our study.
